Fusion Moves for Correlation Clustering

### 2014-11-14 paper/datasets.tex
fixed stuff

%!TEX root = ./egpaper_for_review.tex
\subsection{Datasets}\label{sec:datasets}
\textbf{Social Networks.}\label{sec:nets}
One important application for large scale correlation clustering are social networks.
We consider two of those networks from the Stanford Large Network Dataset Collection\footnote{\url{http://snap.stanford.edu/data/index.html}}.
Both networks are given by weighted directed graphs with edge weights $-1$ and $1$. 
%
The first network is called \emph{Epinions}. 
This is a who-trust-whom online social network of a general consumer review site. 
Each directed edge $a\to b$ indicates that user $a$ trusts  or does not trust user $b$ by a  positive or negative edge-weight, respectively.
The network contains $131828$ nodes and $841372$ edges from which $85.3\%$ are positively weighted.
%
The second network is called \emph{Slashdot}. 
Slashdot is a technology-related news website known for its specific user community. 
In 2002 Slashdot introduced the Slashdot Zoo feature which allows users to tag each other as friend or foe. 
The network was obtained in November 2008 and contains $77350$ nodes and $516575$ edges of which $76.73\%$ are positively weighted.

We consider the problem to cluster these graphs such that positively weighted edges ($E^+_{\to}$) link inside and negatively weighted edges ($E^-_{\to}$) between clusters.
In other words friends and people who trust each other should be in the same segment and foes and non-trusting people in different clusters.
% 
To compensate the high impact of nodes with high degree we can normalize the edge weights such that each person has the same impact on the overall network, by enforcing.
\begin{align}
  \sum_{i\to j \in E_{\to}} |w_{i\to j}| &= 1&\forall i\in V, deg^{\textrm{out}}(i)\geq 1 
\end{align}
We define the following energy function
\begin{align}
 J(y) &= \sum_{i\to j \in E^+_{\to}} y_{ij}\cdot w_{i \to j} +  \sum_{i\to j \in E^-_{\to}} (y_{ij}-1)\cdot w_{i \to j} \nonumber\\
      &= \sum_{ij \in E} y_{ij}\cdot \underbrace{(w_{i \to j}+w_{j \to i})}_{w_{ij}} + \textrm{const}
\end{align}
which is zero if the given partitioning does not violate any relation and larger otherwise.
We name these two datasets \emph{social nets} and \emph{normalized social nets}.

@@

We use the publicly available benchmark instances from~\cite{kappes_2013_benchmark_cvpr,kappes_2014_benchmark_arxiv}.
For 2D images from the Berkeley Segmentation Database~\cite{martin_2001}, we took the segmentation problems called \emph{image-seg}~\cite{andres_2011_iccv,kappes_2013_benchmark_cvpr}.
For 3D volume segmentation we took the models \emph{knott-3d-150},\emph{-300} and \emph{-450} from~\cite{kroeger_2012_eccv,kappes_2014_benchmark_arxiv} as well as the large
instance from the \emph{3d-seg} model~\cite{andres_2011_iccv,kappes_2013_benchmark_cvpr}. These instances have underlying cube sizes of  $150^3$, $300^3$, $450^3$, and $900^3$, respectively.
We also requested larger instances from the authors of~\cite{kroeger_2012_eccv} who kindly provided us the dataset~\emph{knott-3d-550} with cube size  $550^3$.
%More information of the size of instances is given in Tab.~\ref{tab:instance_sizes}.

>>>>

%!TEX root = ./egpaper_for_review.tex
\subsection{Datasets}\label{sec:datasets}
\textbf{Social Networks.}\label{sec:nets}
One important application for large scale correlation clustering are social networks.
We consider two of those networks from the Stanford Large Network Dataset Collection\footnote{\url{http://snap.stanford.edu/data/index.html}}.
Both networks are given by weighted directed graphs with edge weights $-1$ and $+1$. 
%
The first network is called \emph{Epinions}. 
This is a who-trust-whom online social network of a general consumer review site. 
Each directed edge $a\to b$ indicates that user $a$ trusts  or does not trust user $b$ by a  positive or negative edge-weight, respectively.
The network contains $131828$ nodes and $841372$ edges from which $85.3\%$ are positively weighted.
%
The second network is called \emph{Slashdot}. 
Slashdot is a technology-related news website known for its specific user community. 
In 2002 Slashdot introduced the Slashdot Zoo feature which allows users to tag each other as friend or foe. 
The network was obtained in November 2008 and contains $77350$ nodes and $516575$ edges of which $76.73\%$ are positively weighted.

We consider the problem to cluster these graphs such that positively weighted edges ($E^+_{\to}$) link inside and negatively weighted edges ($E^-_{\to}$) between clusters.
In other words friends and people who trust each other should be in the same segment and foes and non-trusting people in different clusters.
% 
To compensate the large impact of nodes with high degree we can normalize the edge weights such that each person has the same impact on the overall network, by enforcing.
\begin{align}
  \sum_{i\to j \in E_{\to}} |w_{i\to j}| &= 1&\forall i\in V, deg^{\textrm{out}}(i)\geq 1 
\end{align}
We define the following energy function
\begin{align}
 J(y) &= \sum_{i\to j \in E^+_{\to}} y_{ij}\cdot w_{i \to j} +  \sum_{i\to j \in E^-_{\to}} (y_{ij}-1)\cdot w_{i \to j} \nonumber\\
      &= \sum_{ij \in E} y_{ij}\cdot \underbrace{(w_{i \to j}+w_{j \to i})}_{w_{ij}} + \textrm{const}
\end{align}
which is zero if the given partitioning does not violate any relation and larger otherwise.
We name these two datasets \emph{social nets} and \emph{normalized social nets}.

@@

We use the publicly available benchmark instances from~\cite{kappes_2013_benchmark_cvpr,kappes_2014_benchmark_arxiv}.
For 2D images from the Berkeley Segmentation Database~\cite{martin_2001}, we took the segmentation problems called \emph{image-seg}~\cite{andres_2011_iccv,kappes_2013_benchmark_cvpr}.
For 3D volume segmentation we use the models \emph{knott-3d-150}, \emph{-300} and \emph{-450} from~\cite{kroeger_2012_eccv,kappes_2014_benchmark_arxiv} as well as the large
instance from the \emph{3d-seg} model~\cite{andres_2011_iccv,kappes_2013_benchmark_cvpr}. These instances have underlying cube sizes of  $150^3$, $300^3$, $450^3$, and $900^3$, respectively.
We also requested larger instances from the authors of~\cite{kroeger_2012_eccv} who kindly provided us the dataset~\emph{knott-3d-550} with cube size  $550^3$.
%More information of the size of instances is given in Tab.~\ref{tab:instance_sizes}.

### 2014-11-14 paper/egpaper_for_review.tex
fixed stuff

\section{Experiments}\label{sec:exp}
In our experiments we compare to the following methods with publicly available implementation.
%
For \textbf{CGC}~\cite{beier_2014_cvpr} we used the publicly available implementation\footnote{\label{opengm-cvpr2014}{\url{github.com/opengm/opengm/tree/cgc-cvpr2014}}} and for \textbf{KL}~\cite{Kernighan-1970} the implementation in OpenGM\footlabel{opengm-master}{\url{github.com/opengm/opengm}}.
For integer multicuts (\textbf{MC-I}) and relaxed multicuts (\textbf{MC-R})~\cite{kappes_2013_arxiv} we use a modified version of
 the publicly available implementation in OpenGM\footref{opengm-master}, as described in Sec.~\ref{sec:tunedMC}.
%
%We realize that their implementation has some limitations on large problems and make the following modifications.
%Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
%Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
%during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
%This procedure is one magnitude faster, but might cause a few extra outer iterations.
%In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
%For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.
%
From the field of data-mining we compare to the PIVOT-algorithm~\cite{Ailon-2008} followed by a round of BOEM~\cite{Gionis-2007} denoted by \textbf{PIVOT-BOEM}.
We use the publicly available code\footnote{\url{http://www.ling.ohio-state.edu/~melsner/resources/correlation-readme.html}}.
As this implementation uses full adjacency matrices it does not scale as well and can not be applied on all datasets.
We also run classical fusion moves~\cite{Lempitsky-2010} (\textbf{Fusion}) and select distinct labels for the two candidate segmentations.
%
As reported in~\cite{beier_2014_cvpr}, CGC is faster and gives better energies than PlanarCC~\cite{yarkony_2012_eccv} and Expand \& Explore~\cite{bagon_2011_arxiv}.
Therefore we exclude those in our experiments.

We compare these methods to the following methods suggested in the present paper.
%
Energy Based Based Hierarchical Clustering (\textbf{EHC}),
as described in Sec.~\ref{sec:ehc}.
%
CGC warm started with the solution from EHC denoted by \textbf{EHC-CGC}.
%
The proposed correlation cluster fusion algorithm with EHC-based and watershed-based proposals
 and MC-I and CGC as subproblem solvers denoted by 
\textbf{CC-Fusion-HC-MC},
\textbf{-HC-CGC},
\textbf{-WS-MC}, and 
\textbf{-WS-CGC}, respectively.
We set the size of the proposals to $10\%$ of the original problem size
and use random edge noise with $\sigma=1.5$.
%
As stopping condition we choose $10^4$ iterations and $100$ iterations with no improvement.

All experiments were run on Intel Core i5-4570 CPUs
with 3.20 GHz, equipped with 32 GB of RAM.
In our evaluation we make no use of multiple threads.
The methods were stopped if they exceed 30 minutes at the next possible interrupt point.

@@

\subsection{Improvements for the Multicut Algorithm}\label{sec:tunedMC} 
When using the  publicly available implementation in OpenGM\footref{opengm-master}
of Kappes~\etal~\cite{kappes_2013_arxiv}, we have noticed that their implementation has some limitations on large problems.
This results in a very slow separation and we make the following modifications.
Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
during the separation procedure. 
In~\cite{kroeger_2012_eccv} this has been used for MC-I only. 
For MC-R this search procedure is not sufficient and needs to be followed by a search for shortest non-chordal paths.
%
Fig.~\ref{fig:at:knott-450b} shows the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
This procedure is one magnitude faster, but might cause a few extra outer iterations.

@@

\subsection{Parameter Choice for CC-Fusion}
Beside the choice of the proposal generator and fusion method CC-Fusion has some more parameters, which need to be set.
The most crucial one is the number of segments in the proposal. For HC we also have to set the noise which is used to generate diversity.
Fig.~\ref{fig:parameterchoice} shows an evaluation of the impact of this parameters for a single instance of knott-3d-450 averaged over several random seeds.
The runtime depends on the number of clusters in the proposal (Fig.~\ref{fig:parameterchoice} left), which controls the size of the auxiliary move problems.
The level of noise has no major impact on the runtime.
The energy of the final solution improves with finer proposals since this increases the search space of the moves.
The level of noise has to be large enough to generate diverse proposals, but not too large as this would lead to proposals with low quality.
As shown in Fig.~\ref{fig:parameterchoice} right the useful parameter set is quite large. This allows us to use the same parameters for \emph{all}
experiments. However we would like to note that in practice we can improve the performance by adjusting these parameters for the specific problem setting.
%
%
\begin{figure}[th]
\centering
\fbox{\includegraphics[width=0.45\columnwidth,trim=3.5cm 8cm 4.5cm 8cm,clip]{images/grid-time2.pdf}}
\fbox{\includegraphics[width=0.45\columnwidth,trim=3.5cm 8cm 4.5cm 8cm,clip]{images/grid-value2.pdf}}
\caption{Empirical evaluation of the impact of noise used for proposal generation and proposal size.
  Proposals with many segments cause longer runtime. Noise seemed not to be a critical parameter but should be selected large enough.
}
\label{fig:parameterchoice}
\end{figure}

@@

\section{Future Work}\label{sec:future}
In future work we would like to investigate much larger problem instances
and interactive correlation clustering, by enforcing
updates only in local regions by appropriate proposals.
%
Within this challenging task we can also enforce additionally must-not-link constraints in eq.~\ref{eq:fusion_move_a}.
If those build a closed surface, the problem decomposes into two independent problems.
%
Overall, a more specific selection of proposals, maybe conditioned on the current solution,
should lead to better proposals and to faster progress of CC-Fusion.
%
Furthermore, we can extend our approach to higher-order correlation clustering~\cite{Kim-2011,kappes_2013_arxiv}
by using higher order subproblems.

>>>>

\section{Experiments}\label{sec:exp}
In our experiments we compare to the following methods with publicly available implementation.
%
For \textbf{CGC}~\cite{beier_2014_cvpr} we used a branch of OpenGM\footnote{\label{opengm-cvpr2014}{\url{github.com/opengm/opengm/tree/cgc-cvpr2014}}} and for \textbf{KL}~\cite{Kernighan-1970} the implementation in OpenGM\footlabel{opengm-master}{\url{github.com/opengm/opengm}}.
For integer multicuts (\textbf{MC-I}) and relaxed multicuts (\textbf{MC-R})~\cite{kappes_2013_arxiv} we modified OpenGM\footref{opengm-master}, as described in Sec.~\ref{sec:tunedMC}.
%
%We realize that their implementation has some limitations on large problems and make the following modifications.
%Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
%Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
%during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
%This procedure is one magnitude faster, but might cause a few extra outer iterations.
%In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
%For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.
%
From the field of data-mining we compare to the PIVOT-algorithm~\cite{Ailon-2008} followed by a round of BOEM~\cite{Gionis-2007} denoted by \textbf{PIVOT-BOEM}\footnote{\url{http://www.ling.ohio-state.edu/~melsner/resources/correlation-readme.html}}.
This implementation uses full adjacency matrices it does not scale and cannot be applied to all datasets.
We also run classical fusion moves~\cite{Lempitsky-2010} (\textbf{Fusion}) and select distinct labels for the two candidate segmentations.
%
According to~\cite{beier_2014_cvpr}, CGC is faster and gives better energies than PlanarCC~\cite{yarkony_2012_eccv} and Expand \& Explore~\cite{bagon_2011_arxiv}.
Therefore we exclude those in our experiments.

We compare all of the above to the following methods suggested in the present paper:
%
Energy Based Hierarchical Clustering (\textbf{EHC}),
as described in Sec.~\ref{sec:ehc}.
%
CGC warm started with the solution from EHC (\textbf{EHC-CGC}).
%
The proposed correlation cluster fusion algorithm with EHC-based and watershed-based proposals
 and MC-I and CGC as subproblem solvers ( 
\textbf{CC-Fusion-HC-MC},
\textbf{-HC-CGC},
\textbf{-WS-MC}, and 
\textbf{-WS-CGC}) respectively.
We set the number of connected components in the proposals to  $10\%$ of the 
number of nodes of
and use random edge noise with $\sigma=1.5$.
%
As stopping condition we choose $10^4$ iterations and $100$ iterations with no improvement.

All experiments were run on Intel Core i5-4570 CPUs
with 3.20 GHz, equipped with 32 GB of RAM.
In our evaluation we make no use of multiple threads.
The methods were stopped once they exceed 30 minutes at the next possible interrupt point.

@@

\subsection{Improvements for the Multicut Algorithm}\label{sec:tunedMC} 
When using the  publicly available implementation in OpenGM\footref{opengm-master}, we have noticed that their implementation has some limitations on large problems.
This results in a very slow separation and we make the following modifications.
Firstly, we used \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm,
which speeds up MC-R.
Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal,instead of searching for the shortest non-chordal path
during the separation procedure. 
In~\cite{kroeger_2012_eccv} this was used for MC-I only. 
For MC-R this search procedure is not sufficient and needs to be followed by a search for shortest non-chordal paths.
%
Fig.~\ref{fig:at:knott-450b} shows the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
This procedure is one magnitude faster, but might cause a few extra outer iterations.

@@

\subsection{Parameter Choice for CC-Fusion}
Beside the choice of the proposal generator and fusion method, CC-Fusion has some more parameters, which need to be set.
The most crucial one is the number of segments in the proposal. For HC we also have to set the noise which is used to generate diversity.
Fig.~\ref{fig:parameterchoice} shows an evaluation of the impact of this parameters for a single instance of knott-3d-450 averaged over several random seeds.
The runtime depends on the number of clusters in the proposal (Fig.~\ref{fig:parameterchoice} left), which controls the size of the auxiliary move problems.
The level of noise has no major impact on the runtime.
The energy of the final solution improves with finer proposals since this increases the search space of the moves.
The level of noise has to be large enough to generate diverse proposals, but not too large as this would lead to proposals with low quality.
As shown in Fig.~\ref{fig:parameterchoice} right the useful parameter set is quite large. This allows us to use the same parameters for \emph{all}
experiments. However we would like to note that in practice we can improve the performance by adjusting these parameters for the specific problem setting.
%
%
\begin{figure}[th]
\centering
\fbox{\includegraphics[width=0.45\columnwidth,trim=3.5cm 8cm 4.5cm 8cm,clip]{images/grid-time2.pdf}}
\fbox{\includegraphics[width=0.45\columnwidth,trim=3.5cm 8cm 4.5cm 8cm,clip]{images/grid-value2.pdf}}
\caption{Empirical evaluation of the impact of noise used for proposal generation and the size of the proposals.
  Proposals with many segments cause longer runtime. Noise seemed not to be a critical parameter but should be selected large enough.
}
\label{fig:parameterchoice}
\end{figure}

@@

\section{Future Work}\label{sec:future}
In future work we would like to investigate much larger problem instances
and interactive correlation clustering, by enforcing
updates only in local regions by appropriate proposals.
%
%Within this challenging task we can also enforce additionally must-not-link constraints in eq.~\ref{eq:fusion_move_a}.
%If those build a closed surface, the problem decomposes into two independent problems.
%
Overall, a more specific selection of proposals, maybe conditioned on the current solution,
should lead to better proposals and to faster progress of CC-Fusion.
%
Furthermore, we can extend our approach to higher-order correlation clustering~\cite{Kim-2011,kappes_2013_arxiv}
by using higher order subproblems.

### 2014-11-14 paper/evaluation.tex
fixed stuff

%!TEX root = ./egpaper_for_review.tex
\subsection{Evaluation}
For the evaluation of the different methods on the datasets introduced in Sec.~\ref{sec:datasets},
we show zoomed anytime plots in  Fig.~\ref{fig:anytime} and variation of information (VOI)~\cite{meila-2003}
and rand index (RI)~\cite{rand-1971} of the final solutions in Tab.~\ref{tab:eval}.
Anytime plots with no zooming and a detailed evaluation are given in the supplementary material. 
%
For \emph{social-nets} CC-Fusion methods provide the best results for the first minutes .
Only CGC and HC-CGC are able to find better solutions after more than 1000 seconds.
MC-I and MC-R can not be applied to such large problems.
We believe that with better proposal generators, which are more suited for such network problems,
we can improve CCFusion. One candidate for such a generator would be a scalable implementation of the 
PIVOT algorithm.
%
For \emph{modularity-clustering} CC-Fusion performs on par to competitive methods, even though CC-Fusion 
and the used parameters have not been designed and chosen for this type of problem.
Especially it does a better job than MC-I.
%
For \emph{image-seg} CC-Fusion is faster than other methods and competitive in terms of energy, VOI, and RI.
Because  the models are designed to have a high boundary recall (oversegmentation), 
classical fusion, which return undersegmentations, has best VOI but worse RI and energy.
Proposals generated by EHC seemed to be a bit better than WS-based ones.
%
For the \emph{knott-datasets}
CC-Fusion-HC-MC and CC-Fusion-WS-MC have a better performance with increasing problem size compared to competitive methods, \cf Fig.~\ref{fig:anytime}(b-e).
Also in terms of VOI and RI they are only slightly worse than the globally optimal solution found by MC-I.
The initialization of CGC by HC denoted by HC-CGC also improves the performance compared to native CGC.
For the largest 3D volume seg-3d, HC-CGC gives the first useful solution, \cf  Fig.~\ref{fig:anytime}(f).
However, after a few minutes  CC-Fusion-HC-MC and CC-Fusion-WS-MC give much better results and are also overall best in terms of energy, VOI, and RI.
%
Pure EHC, Fusion and PIVOT-BOEM do not give useful results on any dataset.

>>>>

%!TEX root = ./egpaper_for_review.tex
\subsection{Evaluation}
For the evaluation of the different methods on the datasets introduced in Sec.~\ref{sec:datasets},
we show zoomed anytime plots in  Fig.~\ref{fig:anytime} and variation of information (VOI)~\cite{meila-2003}
and rand index (RI)~\cite{rand-1971} of the final solutions in Tab.~\ref{tab:eval}.
Anytime plots with no zooming and a detailed evaluation are given in the supplementary material. 
%
For \emph{social-nets} CC-Fusion methods provide the best results for the first minutes .
Only CGC and HC-CGC are able to find better solutions after more than 1000 seconds.
MC-I and MC-R cannot be applied to such large problems.
We believe that with better proposal generators, which are more suited for such network problems,
we can improve CCFusion. One candidate for such a generator would be a scalable implementation of the 
PIVOT algorithm.
%
For \emph{modularity-clustering} CC-Fusion performs on par with competitive methods, even though CC-Fusion 
and the used parameters have not been designed and chosen for this type of problem.
In particular, it does a better job than MC-I.
%
For \emph{image-seg} CC-Fusion is faster than other methods and competitive in terms of energy, VOI, and RI.
Because  the models are designed to have a high boundary recall (oversegmentation), 
classical fusion, which returns undersegmentations, has best VOI but worse RI and energy.
Proposals generated by EHC are a bit better than WS-based ones.
%
For the \emph{knott-datasets}
CC-Fusion-HC-MC and CC-Fusion-WS-MC have a better performance with increasing problem size compared to competitive methods, \cf Fig.~\ref{fig:anytime}(b-e).
Also in terms of VOI and RI they are only slightly worse than the globally optimal solution found by MC-I.
The initialization of CGC by HC, denoted by HC-CGC, also improves the performance compared to native CGC.
For the largest 3D volume seg-3d, HC-CGC gives the first useful solution, \cf  Fig.~\ref{fig:anytime}(f).
However, after a few minutes  CC-Fusion-HC-MC and CC-Fusion-WS-MC give much better results and are also overall best in terms of energy, VOI, and RI.
%
Pure EHC, Fusion and PIVOT-BOEM do not give useful results on any dataset.
